Labeled object representation and manipulation with ITK

### 2007-08-16 article/Article.tex
enhance the article
darcs-hash:20070816100853-2fc9d-c41f329c9d27656df2cbe7d345bccb2bcdd0b46c.gz

The labeled images are often used the represent the connected components of an
image. As connected component, they can be efficiently stored in memory as a set
of lines, using the run-length encoding: a starting point for each line, and the
length of the line on a given dimension (by convention, the dimension 0). The
image the can be seen as a collection of objects.

@@

The {\em itk::LabelCollectionImage} class is in charge of managing the
collection of labeled
objects of the image, as well as storing the metadata associated with the image
like
the spacing, the physical position, etc.

The {\em itk::LabelCollectionImage} provide a part of the API of the {\em
itk::Image} class,
and so can be manipulated as an image\footnote{It doesn't support the itk::Image
iterators though}
in may cases. The performance can be very different however.

@@

No attribute are stored in this class, which can thus be see as the base class
for the objects with some attributes, or which can be used when no attributes
are required.

Several subclasses are provided with that article:
\begin{itemize}
  \item {\em itk::AttributeLabelObject} is able to store a generic attribute. It
is generic in the sense that its type is given in template parameter.
  \item {\em itk::ShapeLabelObject} contains numerous attribute related to the
shape of the labeled object. Computing the values of those attributes does not
require a feaure image.
  \item {\em itk::StatisticsLabelObject} contains numerous statistics about the
grey values of a feature image in the same place than the labeled object.
Computing the values of those attributes does require a feature image.
\end{itemize}

The classes {\em itk::ShapeLabelObject} and {\em itk::StatisticsLabelObject} have been
created to reduce the number of filters made to manipulate the attributes, and to make
the computation of all the set of attributes much efficient. In the early stage of
development, all the attributes were managed as in {\em AttributeLabelObject}, and
a set of 8 classes made to manipulate a single were provided, leading to a huge
number of classes.

@@

@@

The {\em itk::LabelCollectionImage} class provide some methods to fill the
image. However, the most efficient way is to convert a labeled image or a binary
image stored in an {\em itk::Image} to a {\em itk::LabelCollectionImage}, by
using {\em itk::BinaryImageToLabelCollectionImageFilter} or {\em
itk::LabelImageToLabelCollectionImageFilter}.

@@

The label objects produced by those filters have no attribute value set, and
thus the attributes must be valuated. Some filters are provided for the most
common used ones: {\em itk::ShapeLabelCollectionImageFilter} to fill the
attributes of the {\em itk::ShapeLabelObject}s, and {\em
itk::ShapeLabelCollectionImageFilter} to fill the attributes of the {\em
itk::StatisticsLabelObject}s. For the {\em itk::AttributeLabelObject} class or
other classes, the user must set the value by himself, for example by
implementing a subclass of {\em itk::InPlaceLabelCollectionImageFilter}.

@@

Once created and, optionally, valued, several filters are provided to manipulate
the {\em itk::LabelCollectionImage}:

\begin{itemize}
  \item An opening can be performed with the {\em
OpeningLabelCollectionImageFilter} classes. Those classes will remove all the
objects with an attribute value lower or greater than a given value. It's a very
convenient way to enhance a segmentation for example, because we often can use
that way some criteria  which have not been used during the segmentation
procedure, like the size of the object, the mean value of its pixels, etc.
  \item A fixed number of objects can be kept, with the {\em
KeepNObjectsLabelCollectionImageFilter} classes. They are chosen according to
the value of their attribute. The user can choose to keep the ones with the
highest, or with the lowest attribute values.
  \item The objects can be relabel, with the {\em
RelabelLabelCollectionImageFilter} classes. The order of the label is dependant
of the value of the attribute. Again, the user can choose to have the objects
with the highest attribute value in the first labels, or to have the objects
with the lowest attribute values in the first labels.
\end{itemize}

@@

@@

\section{Performance}

TODO

@@

Also, more attributes will be implemented, like the perimeter estimation, the usual moments, etc. It may be useful to implement the most commonly used opening, keep N objects and relabel transforms in a more efficient way, by using an {\em AttributeLabelObject} instead of a {\em ShapeLabelObject} or a {\em StatisticsLabelObject}.

The computation of the moments is partially implemented. The help of someone with a good knownledge of that subject would be very helpful to complete that work. The moments for the binary shape shoudl also be implemented.

>>>>

The labeled images are often used the represent the connected components of an
image. In this contribution, another representation has been chosen.

The objects contained in the image, as connected component, can be efficiently stored in memory as a set
of lines, using the run-length encoding: a starting point for each line, and the
length of the line on a given dimension (by convention, the dimension 0).

The image is a collection of those objects, which also store some values of the image, like its size, its spacing, etc.

@@

The {\em itk::LabelCollectionImage} class is in charge of managing the
collection of labeled
objects of the image, as well as storing the metadata associated with the image
like
the spacing, the physical position - all the metadata found in {\em itk::Image}.

The {\em itk::LabelCollectionImage} provide a part of the API of the {\em
itk::Image} class, and so can be manipulated as an image\footnote{It doesn't
support the itk::Image iterators though} in many cases. The performance can be
very different however, because of the very different data structure used.

The {\em itk::LabelCollectionImage} is a templated class, which take a single
parameter: the type of {\em labeled object} stored by that class. The dimension
of the image is took from the {\em labeled object} class, and thus don't need to
be defined as template parameter of that class. The pixel type of the image
also comes from the {\em labeled object} class.

@@

No attribute are stored in this class, which can thus be seen as the base class
for the objects with attributes, or which can be used when no attributes
are required.

The {\em itk::LabelObject} class is templated and takes to required template parameters:
\begin{itemize}
  \item the type of the label,
  \item the dimension of the image.
\end{itemize}

Several subclasses are provided with that contribution, to cover the most common
usages of the labeled objects manipulation:
\begin{itemize}
  \item {\em itk::AttributeLabelObject} is able to store a generic attribute. It
is generic in the sense that its type is given in template parameter.
  \item {\em itk::ShapeLabelObject} contains numerous attribute related to the
shape of the labeled object. Computing the values of those attributes does not
require a feaure image.
  \item {\em itk::StatisticsLabelObject} contains numerous statistics about the
grey values of a feature image in the same place than the labeled object.
Computing the values of those attributes {\em does} require a feature image.
\end{itemize}

The classes {\em itk::ShapeLabelObject} and {\em itk::StatisticsLabelObject} have been
created to reduce the number of filters made to manipulate the attributes, and to make
the computation of all the set of attributes much efficient. In the early stage of
development, all the attributes were managed as in {\em AttributeLabelObject}, and
a set of 8 classes made to manipulate a single attribute were provided, leading to a huge
number of classes.

@@

Both {\em itk::ShapeLabelObject} and {\em itk::StatisticsLabelObject} are templated classes.
They take the same template parameters than the {\em itk::LabelObject} class.
The 2 first template parameters of the {\em itk::AttributeLabelObject} class or the same
than the {\em itk::LabelObject} class. The third one is the attribute type.

@@

The {\em itk::LabelCollectionImage} class provide some methods to fill the
image "by hand", like the usual {\em SetPixel()} method. However, the most efficient way is to convert a labeled image or a binary
image stored in an {\em itk::Image} to a {\em itk::LabelCollectionImage}, by
using {\em itk::BinaryImageToLabelCollectionImageFilter} or {\em
itk::LabelImageToLabelCollectionImageFilter}.

@@

The label objects produced by those filters have no attribute value set, and
thus, the attributes must be valuated. Some filters are provided for the most
common used ones:

\begin{itemize}
  \item {\em itk::ShapeLabelCollectionImageFilter} to fill the
attributes of the {\em itk::ShapeLabelObject}s, 
  \item and {\em itk::ShapeLabelCollectionImageFilter} to fill the attributes of the {\em
itk::StatisticsLabelObject}s.
\end{itemize}

For the {\em itk::AttributeLabelObject} class or
other classes, the user must set the value by himself, for example by
implementing a subclass of {\em itk::InPlaceLabelCollectionImageFilter}.

@@

Once created and, optionally, valuated, several filters are provided to manipulate
the {\em itk::LabelCollectionImage}:

\begin{itemize}
  \item An opening can be performed with the {\em
OpeningLabelCollectionImageFilter} classes. Those classes will remove all the
objects with an attribute value lower or greater than a given value.
Because we often can use some criteria  which have not been used during the segmentation
procedure, like the size of the object, the mean value of its pixels, etc., the attribute
opening is often a very efficient way to enhance a segmentation. For example, after
a thresholding of a grayscale image, the objects too small or too beg to be of interest
can be removed that way.

  \item A fixed number of objects can be kept, with the {\em
KeepNObjectsLabelCollectionImageFilter} classes. They are chosen according to
the value of their attribute. The user can choose to keep the ones with the
highest, or with the lowest attribute values.

  \item The objects can be relabel, with the {\em
RelabelLabelCollectionImageFilter} classes. The order of the label is dependant
of the value of the attribute. Again, the user can choose to have the objects
with the highest attribute value in the first labels, or to have the objects
with the lowest attribute values in the first labels.
\end{itemize}

@@

\subsection{Reading attribute values}

In that example, we will read a binary image, and get some of attributes
about the obejcts contained in that image.
The source code is available in the file {\em attribute\_values.cxx}.

First include the classes we'll use
\small \begin{verbatim}
#include "itkImageFileReader.h"
#include "itkShapeLabelObject.h"
#include "itkLabelCollectionImage.h"
#include "itkBinaryImageToLabelCollectionImageFilter.h"
#include "itkShapeLabelCollectionImageFilter.h"

int main(int, char * argv[])
{
  const int dim = 2;
\end{verbatim} \normalsize
then declare the type of the input image
\small \begin{verbatim}
  typedef unsigned char PixelType;
  typedef itk::Image< PixelType, dim >    ImageType;

\end{verbatim} \normalsize
read the input image
\small \begin{verbatim}
  typedef itk::ImageFileReader< ImageType > ReaderType;
  ReaderType::Pointer reader = ReaderType::New();
  reader->SetFileName( argv[1] );

\end{verbatim} \normalsize
define the object type. Here the ShapeLabelObject type
is chosen in order to read some attribute related to the shape
of the objects (by opposition to the content of the object, with
the StatisticsLabelObejct).
\small \begin{verbatim}
  typedef unsigned long LabelType;
  typedef itk::ShapeLabelObject< LabelType, dim > LabelObjectType;
  typedef itk::LabelCollectionImage< LabelObjectType > LabelCollectionType;

\end{verbatim} \normalsize
convert the image in a collection of objects
\small \begin{verbatim}
  typedef itk::BinaryImageToLabelCollectionImageFilter< ImageType, LabelCollectionType > ConverterType;
  ConverterType::Pointer converter = ConverterType::New();
  converter->SetInput( reader->GetOutput() );
  converter->SetForegroundValue( 200 );

\end{verbatim} \normalsize
and valuate the attributes with the dedicated filter: ShapeLabelCollectionImageFilter
\small \begin{verbatim}
  typedef itk::ShapeLabelCollectionImageFilter< LabelCollectionType > ShapeFilterType;
  ShapeFilterType::Pointer shape = ShapeFilterType::New();
  shape->SetInput( converter->GetOutput() );

\end{verbatim} \normalsize
update the shape filter, so its output will be up to date
\small \begin{verbatim}
  shape->Update();

\end{verbatim} \normalsize
then we can read the attribute values we're interested in. {\em BinaryImageToLabelCollectionImageFilter}
produces consecutives labels, so a simple {for} loop will do the job.
\small \begin{verbatim}
  LabelCollectionType::Pointer collection = shape->GetOutput();
  for( int label=1; label<collection->GetNumberOfObjects(); label++ )
    {
    LabelObjectType::Pointer labelObject = collection->GetLabelObject( label );
    std::cout << label << "\t" << labelObject->GetPhysicalSize() << "\t" << labelObject->GetCentroid() << std::endl;
    }

@@

% \section{Performance}
% 
% TODO

@@

Also, more attributes will be implemented, like the perimeter estimation, the usual image moments, etc. It may be useful to implement the most commonly used opening, keep N objects and relabel transforms in a more efficient way, by using an {\em AttributeLabelObject} instead of a {\em ShapeLabelObject} or a {\em StatisticsLabelObject}.

The computation of the moments is partially implemented. The help of someone with a good knownledge of that subject would be very helpful to complete that work. The moments for the binary shape should also be implemented.
